Tactics and Marks in Banach Mazur Games

### 2018-02-23 postdoctoral/k-markov-tactic-banach-mazur.tex
almost finish poset banach mazur game result

  \begin{definition}
    Let \(\mb P\) be partially ordered by \(\leq\).
    Let \(\mb P^{\downarrow}=\{f\in\mb P^\omega : f(n)\geq f(n+1)\}\).
    Then for \(f,g\in\mb P^\downarrow\), we say that \(f,g\) zip into each
    other if for all \(m<\omega\) there exists \(n<\omega\) such that
    \(f(m)\geq g(n)\) and \(g(m)\geq f(n)\).
  \end{definition}

  \begin{definition}
    \(\bmPoGame{\mb P}{W}\) is a game defined for all non-empty partial orders
    \(\mb P\) and all subsets \(W\subseteq\mb P^\downarrow\).
    During round \(0\), \(\plI\) chooses \(a_0\in\mb P\),
    and then \(\plII\) chooses \(b_0\leq a_0\); during around \(n+1\),
    \(\plI\) chooses \(a_{n+1}\leq b_n\), and then \(\plII\) chooses
    \(b_{n+1}\leq a_{n+1}\). \(\plII\) wins this game if
    \(\<a_0,a_1,\dots\>\in W\).
  \end{definition}

  \begin{theorem}
    Let \(W\subseteq\mb P^\downarrow\) be closed under zipping.
    \(\plII\markwin\bmPoGame{\mb P}{W}\) if and only if
    \(\plII\tactwin\bmPoGame{\mb P}{W}\).
  \end{theorem}

@@

    Suppose \(\sigma\) is legally attacked by \(a\in\mb P^\omega\).
    For \(n\leq\omega\), if \(a(n)\) is long, then \(a(n)\geq_n l(a(n))\).
    Therefore,
    \[
      a(n)
        \geq
      s_n(a(n))
        \geq
      \tau(s_n(a(n)),n+1)
        \geq
      l(a(n))
        =
      \sigma(a(n))
        \geq
      a(n+1)
    .\]
    Thus if \(a(n)\) is long for \(n<\omega\), it follows that
    \(c\in\mb P^\downarrow\) defined by \(c(n)=s_n(a(n))\)
    is a legal attack against \(\tau\). Since \(\tau\) is winning,
    \(c\in W\), and since \(c\) zips into \(a\),
    \(a\in W\) as well.

@@

    Thus for \(n<\omega\),
    \[
      b(n)
        \geq
      \tau^{n}(b(n))
        \geq
      \tau(\tau^{n}(b(n)),n+1)
        \geq
      \tau^{N(b(n))+1}(b(n))
        =
      \sigma(b(n))
        \geq
      b(n+1)
    .\]
    As a result, \(c\in\mb P^\downarrow\) defined by \(c(n)=\tau^n(b(n))\)
    is a legal attack against the winning strategy \(\tau\). Therefore
    \(c\in W\), and since \(c\) zips into \(b\) and \(a\), we conclude
    \(a\in W\).
  \end{proof}

  \begin{observation}
    When \(\mb P=T(X)\setminus\{\emptyset\}\) is ordered by set-inclusion
    and \(W=\{U\in\mb P^\downarrow:\bigcap_{n<\omega}U(n)\not=\emptyset\}\),
    then \(\bmPoGame{\mb P}{W}\) is exactly the topological Banach Mazur game
    \(\bmGame{X}\). Note \(W\) is closed under zipping.
  \end{observation}

@@

    % \begin{proposition}
    %   Suppose that for every finite collection \(\{\mb P_i:i<n\}\)
    %   such that \(\mb P=\bigcup_{i<n}\mb P_i\),
    %   some \(\mb P_i\) contains two incompatible points.
    %   Then \(\mb P\) contains an infinite antichain.
    % \end{proposition}
    %
    % \begin{proof}
    %   Suppose \(a_i\in\mb P_i\) have been defined for \(i<n\) such that
    %   \(a_i\not\in\bigcup_{j<i}\mb P_j\) and
    %   each \(\mb P_i\) is pairwise-compatible. It follows that
    %   \(\mb P\not=\bigcup_{i<n}\mb P_i\), so choose
    %   \(a_n\not\in\bigcup_{i<n}\mb P_i\) and let \(\mb P_n\) be a maximal family
    %   containing \(a_n\) such that \(\mb P_n\) is pairwise-compatible.
    %
    %   It follows that \(A=\{a_n:n<\omega\}\) is an antichain, since for
    %   every pair \(a_i,a_j\in A\) with \(i<j\), \(a_j\not\in\mb P_i\), so
    %   \(a_j\) is not compatible with something in \(\mb P_i\)...
    % \end{proof}

@@

\begin{definition}
  For every partial order \(\mb P\) and compatible \(p,q\in\mb P\),
  let \(p\wedge q\) satisfy \(p\wedge q\leq p,q\).
\end{definition}

  \begin{claim}
    \(\mb P\) contains no infinite antichains if and only if every
    antichain in \(\mb P\) is of size \(n\) or less for some \(n<\omega\).
  \end{claim}

  \begin{proof}
    MAYBE? Apparently true for \(\mb P=\tau\setminus\{\emptyset\}\) due to
    Lemma 2.10 of \cite{comfort_negrepontis_1982}.
  \end{proof}

  \begin{proposition}
    Let \(W\subseteq\mb P^\downarrow\) be closed under zipping.
    Suppose every antichain in \(\mb P\) is of size \(n<\omega\) or less,
    and \(\plII\win\bmPoGame{\mb P}{W}\). Then
    \(\plII\autowin\bmPoGame{\mb P}{W}\) (i.e. \(\plII\) wins every
    play of \(\bmPoGame{\mb P}{W}\), i.e. \(W=\mb P^\downarrow\)).
  \end{proposition}

@@

  \begin{lemma}
    Let \(W\subseteq\mb P^\downarrow\) be closed under zipping.
    Suppose that for every \(p\in\mb P\), there exists an infinite
    antichain \(A_p=\{a_p(n):n<\omega\}\subseteq\{q\in\mb P:q\leq p\}\).
    Then \(\plII\kmarkwin{(k+2)}\bmPoGame{\mb P}{W}\) if and only if
    \(\plII\ktactwin{(k+2)}\bmPoGame{\mb P}{W}\).
  \end{lemma}

@@

    Define \(g\in\mb P^\downarrow\) by \(g(0)=f(0)\) and
    \(g(j+1)=a_{f(j+1)}(j+1)\).
    Reviewing the above, the reader may confirm that we have shown for
    \(n<k+2\)
    \[
      f(n+1)
        \leq
      \tau(f\rest(n+1))
        =
      \sigma(g\rest(n+1),n+1)
        \leq
      g(n)
        \leq
      f(n)
    \]
    and for \(n\geq k+2\)
    \[
      f(n+1)
        \leq
      \tau(f\rest(n+1)\last(k+2))
        =
      \sigma(g\rest(n+1)\last(k+2),n+1)
        \leq
      g(n)
        \leq
      f(n)
    \]
    Thus \(g\) is a legal attack against \(\sigma\), and since \(\sigma\)
    is winning, \(g\in W\). Since \(W\) is closed under zipping,
    \(f\in W\), and therefore \(\tau\) is also winning.
  \end{proof}

>>>>

  \begin{definition}
    Let \(\mb P\) be partially ordered by \(\leq\).
    Let \(\mb P^{\omega,\downarrow}=\{f\in\mb P^\omega : f(n)\geq f(n+1)\}\).
    Then for \(f,g\in\mb P^{\omega,\downarrow}\), we say that \(f,g\) zip into each
    other if for all \(m<\omega\) there exists \(n<\omega\) such that
    \(f(m)\geq g(n)\) and \(g(m)\geq f(n)\).
  \end{definition}

  \begin{definition}
    \(\bmPoGame{\mb P}{W}\) is a game defined for all non-empty partial orders
    \(\mb P\) and all subsets \(W\subseteq\mb P^{\omega,\downarrow}\).
    During round \(0\), \(\plI\) chooses \(a_0\in\mb P\),
    and then \(\plII\) chooses \(b_0\leq a_0\); during around \(n+1\),
    \(\plI\) chooses \(a_{n+1}\leq b_n\), and then \(\plII\) chooses
    \(b_{n+1}\leq a_{n+1}\). \(\plII\) wins this game if
    \(\<a_0,a_1,\dots\>\in W\).
  \end{definition}

  \begin{theorem}
    Let \(W\subseteq\mb P^{\omega,\downarrow}\) be closed under zipping.
    \(\plII\markwin\bmPoGame{\mb P}{W}\) if and only if
    \(\plII\tactwin\bmPoGame{\mb P}{W}\).
  \end{theorem}

@@

    Suppose \(\sigma\) is legally attacked by \(a\in\mb P^\omega\).
    For \(n\leq\omega\), if \(a(n)\) is long, then \(a(n)\geq_n l(a(n))\).
    Therefore,
    \[
      a(n)
        \geq
      s_n(a(n))
        \geq
      \tau(s_n(a(n)),n+1)
        \geq
      l(a(n))
        =
      \sigma(a(n))
        \geq
      a(n+1)
    .\]
    Thus if \(a(n)\) is long for \(n<\omega\), it follows that
    \(c\in\mb P^{\omega,\downarrow}\) defined by \(c(n)=s_n(a(n))\)
    is a legal attack against \(\tau\). Since \(\tau\) is winning,
    \(c\in W\), and since \(c\) zips into \(a\),
    \(a\in W\) as well.

@@

    Thus for \(n<\omega\),
    \[
      b(n)
        \geq
      \tau^{n}(b(n))
        \geq
      \tau(\tau^{n}(b(n)),n+1)
        \geq
      \tau^{N(b(n))+1}(b(n))
        =
      \sigma(b(n))
        \geq
      b(n+1)
    .\]
    As a result, \(c\in\mb P^{\omega,\downarrow}\) defined by \(c(n)=\tau^n(b(n))\)
    is a legal attack against the winning strategy \(\tau\). Therefore
    \(c\in W\), and since \(c\) zips into \(b\) and \(a\), we conclude
    \(a\in W\).
  \end{proof}

  \begin{observation}
    When \(\mb P=T(X)\setminus\{\emptyset\}\) is ordered by set-inclusion
    and \(W=\{U\in\mb P^{\omega,\downarrow}:\bigcap_{n<\omega}U(n)\not=\emptyset\}\),
    then \(\bmPoGame{\mb P}{W}\) is exactly the topological Banach Mazur game
    \(\bmGame{X}\). Note \(W\) is closed under zipping.
  \end{observation}

@@

@@

  \begin{definition}
    For every partial order \(\mb P\) and compatible \(p,q\in\mb P\),
    write \(p\not\perp q\) and let \(p\wedge q\) satisfy \(p\wedge q\leq p,q\).
    If \(p,q\) are incompatible, write \(p\perp q\).
  \end{definition}

  \begin{definition}
    For every partial order \(\mb P\) and compatible \(p\in\mb P\),
    let \(p^\downarrow=\{q\in\mb P:q\leq p\}\).
\end{definition}

  \begin{lemma}
    \(\mb P\) contains no infinite antichains if and only if every
    antichain in \(\mb P\) is of size \(n\) or less for some \(n<\omega\).
  \end{lemma}

  \begin{proof}
    This was shown to be true for \(\mb P=\tau\setminus\{\emptyset\}\) in
    Lemma 2.10 of \cite{comfort_negrepontis_1982}. It's likely known
    for general \(\mb P\), but I can't find a citation, so let's roll
    our own proof here. Assume \(\mb P\) has antichains of size \(n\)
    for all \(n<\omega\).

    Say \(p\in\mb P\) is bad if there exists \(r_p\leq p\) such that
    \(r_p^\downarrow\) is pairwise compatible. Let
    \(\mb P_{bad}\) collect all bad points in \(\mb P\), and say
    \(p\sim q\) for \(p,q\in\mb P_{bad}\) if \(r_p\not\perp r_q\).
    This is obviously symmetric and reflexive, and if we assume
    \(p\sim q,q\sim t\), then let \(s_p=r_p\wedge r_q\) and
    \(s_r=r_q\wedge r_t\). Since \(r_q^\downarrow\in\mb P_{bad}\),
    \(s_p\not\perp s_r\), so \(r_p\not\perp r_t\) and thus
    \(p\sim t\). Thus \(\sim\) is an equivalence relation.

    If \(\mb P_{bad}/\sim\) is infinite, we may choose \(p_i\in\mb P_{bad}\)
    such that \(p_i\not\sim p_j\) for \(i<j<\omega\). Thus
    \(r_{p_i}\perp r_{p_j}\) for \(i<j<\omega\), giving us
    an infinite antichain \(\{r_{p_i}:i<\omega\}\).

    Otherwise \(|\mb P_{bad}/\sim|=n<\omega\), and choose an antichain
    \(\{p_i:i\leq n\}\) in \(\mb P\). If \(\{p_i:i<n\}\subseteq\mb P_{bad}\),
    \(p_i\perp p_j\) implies \(r_{p_i}\perp r_{p_j}\)
    and \(p_i\not\sim p_j\) for all
    \(i\leq n\). Thus \(\mb P_{bad}=\bigcup_{i<n}\overset{\sim}{p_i}\),
    and \(p_n\not\in\mb P_{bad}\).

    So we've found \(b_0\in\mb P\setminus\mb P_{bad}\).
    Given \(b_n\in\mb P\setminus\mb P_{bad}\), we may choose
    \(a_n,b_{n+1}\leq b_n\) such that \(a_n\perp b_{n+1}\).
    Thus by construction, \(a_n\perp a_{m+1}\) for all
    \(n\leq m<\omega\). Therefore
    \(\{a_n:n<\omega\}\) is an antichain.
  \end{proof}

  \begin{proposition}
    Let \(W\subseteq\mb P^{\omega,\downarrow}\) be closed under zipping.
    Suppose every antichain in \(\mb P\) is of size \(n<\omega\) or less,
    and \(\plII\win\bmPoGame{\mb P}{W}\). Then
    \(\plII\autowin\bmPoGame{\mb P}{W}\) (i.e. \(\plII\) wins every
    play of \(\bmPoGame{\mb P}{W}\), i.e. \(W=\mb P^{\omega,\downarrow}\)).
  \end{proposition}

@@

  \begin{lemma}
    Let \(W\subseteq\mb P^{\omega,\downarrow}\) be closed under zipping.
    Suppose that for every \(p\in\mb P\), there exists an infinite
    antichain \(A_p=\{a_p(n):n<\omega\}\subseteq\{q\in\mb P:q\leq p\}\).
    Then \(\plII\kmarkwin{(k+2)}\bmPoGame{\mb P}{W}\) if and only if
    \(\plII\ktactwin{(k+2)}\bmPoGame{\mb P}{W}\).
  \end{lemma}

@@

    Define \(g\in\mb P^{\omega,\downarrow}\) by \(g(0)=f(0)\) and
    \(g(j+1)=a_{f(j+1)}(j+1)\).
    Reviewing the above, the reader may confirm that we have shown for
    \(n<k+2\)
    \[
      f(n+1)
        \leq
      \tau(f\rest(n+1))
        =
      \sigma(g\rest(n+1),n+1)
        \leq
      g(n)
        \leq
      f(n)
    \]
    and for \(n\geq k+2\)
    \[
      f(n+1)
        \leq
      \tau(f\rest(n+1)\last(k+2))
        =
      \sigma(g\rest(n+1)\last(k+2),n+1)
        \leq
      g(n)
        \leq
      f(n)
    \]
    Thus \(g\) is a legal attack against \(\sigma\), and since \(\sigma\)
    is winning, \(g\in W\). Since \(W\) is closed under zipping,
    \(f\in W\), and therefore \(\tau\) is also winning.
  \end{proof}
